Decoding the Amplitude-Modulated Part the DCF77 Longwave Time Signal by using the Goertzel Algorithm

### 2024-08-22 paper/main.tex
Housekeeping, restructuring and simplifications. Fixed typos

long where it has a value to zero.

@@

By using this scheme, a dataframe consisting out \SI{59}{\Bit}, starting at second one, can be transmitted every minute.

@@

With (\ref{eqn:fs}), a table containing possible ranges for $f_{s}$ can be created.

@@

At $m = 32$, the condition of (\ref{eqn:fs_b}) is violated and at $m = 32$ the condition of (\ref{eqn:fs}), making $m = 31$ the last valid range.
Choosing a sampling rate can now be done from this table while.

@@

distance between the lower and the upper boundaty of the sampling rate.
This then leads to higher requirements regarding the correctness of the sampling rate.
A choice providing a good tradeoff between these effects is $m = 7$ with $f_{s} = \SI{24}{\kilo\hertz}$, because it offers a distance of $\SI{2647}{\hertz}$
the upper and lower boundary with the sampling frequency lying almost in the middle \cite{b10}.

@@

caused by the convolution of the initial spectrum with the dirac delta impulse, located at $-3 \cdot f_{s}$. 

@@

\subsection{The Goertzel Algorithm}
The standard approach to analyze a spectrum in the digital domain after sampling is the well-known Discrete Fourier Transform (DFT), implemented by using
the Fast Fourier Transform (FFT). This yields an analysis of the complete spectrum, producing $\frac{N}{2} + 1$ relevant results for equally distributed frequencies
("frequency bins") within range from \SI{0}{\hertz} to $\frac{f_{s}}{2}$ where $N$ is the number of samples and $f_{s}$ the sampling frequency \cite{b6}.
The signal strength, and therefore the binary time signal of the amplitude-modulated part, can then be examined by observing the value of the curresponding frequency bin
over time. Because of only one frequency bin being relevant, it is obvious that a full FFT produces a lot of overhead and redundant calculation operations in this use-case. 
This overhead can be avoided by using the Goertzel algorithm instead of the FFT, which is able to calculate the value of one specific frequency bin with a smaller
amount of needed calculations.
It was invented in 1958 by Gerald Goertzel and since then is most commonly used in the detection tones employed in Dual-tone Multi-frequency (DTMF) signaling \cite{b7}.
The working principle of the algorithm can be percieved as a second-order Infinite Impulse Response (IIR) filter where the results $y[n-1]$ and $y[n-2]$ of the output $y[n]$
are stored in states $Q_{1}$ and $Q_{2}$, so they are still available after the execution of the filter \cite{b8}.
\begin{equation}
    y[n] = x[n] + 2 \cdot cos(\omega_{g}) \cdot y[n-1] - y[n-2]
\end{equation}
\begin{figure}[h]
    \centering
    \resizebox{0.4\textwidth}{!}{
        \begin{tikzpicture}[auto, node distance=2cm]
            %Defining a cross set for the multiplication elements
            \tikzset{cross/.style={path picture={ 
            \draw[black]
            (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);
            }}}

@@

            \draw[->] (x.east) -- (sum1.west);
            \draw[->] (sum1.east) -- (y.west);
            \draw[->] (sum2.north) -- (sum1.south);
            \draw[->] (mult1.west) -- (sum2.east);
            \draw[->] (mult2.west) -- (0, -4) -- (sum2.south);
            \draw[->] (4, 0) -- (del1.north);
            \draw[->] (del1.south) -- (del2.north);
            \draw[->] (del2.south) -- (4,-4) -- (mult2.east);
            \draw[->] (4,-2) -- (mult1.east);
            \draw[->] (2, -1.25) -- (mult1.north);
            \draw[->] (2, -3.25) -- (mult2.north);
            \draw[->] (4, -2) -- (st1.west);
            \draw[->] (4, -4) -- (st2.west);
            \node at (2, -1) {$2 \cdot cos(\omega_{g})$};
            \node at (2, -3) {$-1$};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,0) {};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,-2) {};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,-4) {};
        \end{tikzpicture}
    }
    \caption{IIR representation of the Goertzel algorithm without the final calculation of the squared magnitude}
    \label{fig:goertzel_iir}
\end{figure}
\FloatBarrier\noindent
The system takes the input sequence $x[n]$ with length $N$ and is executed exactly $N$ times.
After the execution, the squared magnitude of $f_{a}$, the frequency to be analyzed, inside $x[n]$ can be calculated using the two states $Q_{1}$ and $Q_{2}$:
\begin{equation}
    |X(f_{a})|^2 = Q_{1}^2 + Q_{2}^2 - Q_{1} \cdot Q_{2} \cdot 2 \cdot \cos(\omega_{g})
\end{equation}
The constant $2 \cdot cos(\omega_{g})$, often called the "Goertzel coefficient", is calculated before the execution and used to adapt the algorithm to the frequency to be analyzed.
The Value of $\omega_{g}$ itself is calculated based on a integer number $k$ and $N$.
The integer number $k$ is a rounded fraction based on $N$, the frequency to be analyzed $f_{a}$, and the the sampling frequency $f_{s}$.
\begin{equation}
    k = \left\lfloor \frac{N \cdot f_{a}}{f_{s}} \right\rceil
\end{equation}
\begin{equation}
    w_g = \frac{2\pi \cdot k}{N}
\end{equation}

\par
Just as with the conventional FFT, the amount of needed calculations for executing the Goertzel algorithm is based on the number of samples.
In total, $N + 1$ multiplications and $2 \cdot N + 2$ additions are needed, which can be seen from the formulas for the IIR-part and the final calculation of the squared magnitude.
For example, a length of $N = 512$ requires in $513$ multiplications and $1026$ additions to calculate the squared magnitude of the desired frequency.
In comparison, the typical radix-2 FFT, needs $\frac{N}{2} \cdot \log_{2}(N)$ complex multiplications and $N \cdot \log_{2}(N)$ complex additions. With $N = 512$ this would result in $2034$ complex multiplications and $4608$ complex additions.
\par
With this simple example, it is already visible that the Goertzel algorithm can reduce the amount of needed resources compared to a full FFT, when only one or a small amount of frequency bins should be analyzed.
It is therefore used to observe the DCF77 signal's amplitude over time by being executed periodically.
As the gap of the encoded symbols is eiter \SI{100}{\milli\second} or \SI{200}{\milli\second} long, the duration of the analyzation set to \SI{10}{\milli\second} to provide a good tradeoff between resolution in time and a longest possible analyzation time
to leverage the DFT gain.
The square roots of the results are taken to finally calculate the magnitude of the DCF77 signal's amplitude and then passed to a detector to obtain the binary time signal.
Because of the duration of the analyzation being equal to \SI{10}{\milli\second}, the sampling rate after the Goertzel algorithm is equal to:
\begin{equation}
    f_{s,g} = \frac{1}{\SI{10}{\milli\second}} = \SI{100}{\hertz}
\end{equation}

\subsection{The detector}
The output of the Goertzel algorithm is already following the waveform of the binary time signal but still includes fluctuations due to noise and fading.
To reconstruct the original waveform of the binary time signal, it is processed inside a detector.
As the values of the Goertzel algorithm will follow the pattern of the signal's amplitude being switched between \SI{15}{\percent} and \SI{100}, the detection can be done by comparing the current value to a threshold.
Because of the Goertzel algorithm delivering absolute values, whose magnitude depend on the current receiving situation, a Exponential Moving Average is calculated and used to normalize the values by using (\ref{eqn:ema}).
\begin{equation}
    y[t] = \alpha \cdot x[t] + (1 - \alpha) \cdot y[t-1]
    \label{eqn:ema}
\end{equation}
This part can also be represented as an IIR-structure:
\begin{figure}[h]
    \centering
    \resizebox{0.4\textwidth}{!}{
        \begin{tikzpicture}[auto, node distance=2cm]
            %Defining a cross set for the multiplication elements
            \tikzset{cross/.style={path picture={ 
            \draw[black]
            (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);
            }}}

@@

            \draw[->] (x.east) -- (mult1.west);
            \draw[->] (mult1.east) -- (sum1.west);
            \draw[->] (sum1.east) -- (y.west);
            \draw[->] (mult2.west) -- (1.5, -2) -- (sum1.south);
            \draw[->] (del1.south) -- (4.5, -2) -- (mult2.east);
            \draw[->] (4.5, 0) -- (del1.north);
            \draw[->] (0, 0.75) -- (mult1.north);
            \draw[->] (3, -1.25) -- (mult2.north);
            \node at (0, 1) {$\alpha$};
            \node at (3, -1) {$1 - \alpha$};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4.5,0) {};
        \end{tikzpicture}
    }
    \caption{IIR representation of the Exponential Moving Average}
    \label{fig:ema_iir}
\end{figure}
\FloatBarrier\noindent
The value of the smoothing factor $\alpha$ was calculated to take the samples of the last \SI{10}{\second} into account:
\begin{equation}
    \begin{split}
        \alpha &= \frac{2}{N + 1} = \frac{2}{\SI{10}{\second} \cdot f_{s,g} + 1}\\
        &= \frac{2}{\SI{10}{\second} \cdot \SI{100}{\hertz} + 1} \approx 0.001998
    \end{split}
\end{equation}
This provides a stable and slowly changing average, as it is putting more weight on the past samples and leaves minimal influence to the most recent sample.
The decision is then done by comparing the current sample to a threshold calculated by multiplying the current value of the average with $0.575$ (the middle bewteen \SI{15}{\percent} and \SI{100}{\percent}) and a scaling factor of $\frac{1}{0.8725}$.
The scaling factor is needed as the average percentage of time where the signal where is equal to \SI{100}{percent} is \SI{87.25}{\percent} (assuming an equal distribution between symbols representing a 1 and a 0).

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a 
separate text file. Complete all content and organizational editing before 
formatting. Please note sections \ref{AA} to \ref{FAT} below for more information on 
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been 
formatted and styled. Do not number text heads---{\LaTeX} will do that 
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have been defined in the abstract. Abbreviations such as 
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}\label{AAA}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}\label{ITH}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}\label{FAT}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab}
\end{center}
\end{table}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

>>>>

long where it has a value equal to 0.

@@

By using this scheme, a dataframe consisting out \SI{59}{\Bit}, starting at second 1, can be transmitted every minute.

@@

With \eqref{eqn:fs}, a table containing possible ranges for $f_{s}$ can be created.

@@

At $m = 32$, the condition of \eqref{eqn:fs_b} is violated and at $m = 32$ the condition of \eqref{eqn:fs}, making $m = 31$ the last valid range.
Choosing a sampling rate can now be done with this table.

@@

distance between the lower and the upper limits of the sampling rate.
This then leads to higher requirements regarding the adherence to the exact sampling rate.
A choice providing a good tradeoff between these effects $f_{s} = \SI{24}{\kilo\hertz}$ at $m = 7$, because it offers a distance of $\SI{2647}{\hertz}$ between the upper and lower limits with the sampling frequency lying almost in the middle \cite{b10}.

@@

caused by the convolution of the initial spectrum with the dirac delta impulse located at $-3 \cdot f_{s}$. 

@@

\subsection{The Goertzel Algorithm}
The standard approach to analyze a spectrum in the digital domain after sampling is the well-known Discrete Fourier Transform (DFT), implemented by using
the Fast Fourier Transform (FFT). This yields an analysis of the complete spectrum, producing $\frac{N}{2} + 1$ relevant results for equally distributed frequencies
("frequency bins") within a range from \SI{0}{\hertz} to $\frac{f_{s}}{2}$ where $N$ is the number of samples and $f_{s}$ is the sampling frequency \cite{b6}.
The signal strength, and therefore the binary time signal of the amplitude-modulated part, can then be examined by observing the magnitude of the curresponding frequency bin
over time. Because of only one frequency bin being relevant, it is obvious that a full FFT produces a lot of overhead and redundant calculation operations in this use-case. 
This overhead can be avoided by using the Goertzel algorithm instead of the FFT, which is able to calculate the value of one specific frequency bin with a smaller
amount of needed calculations.
It was invented in 1958 by Gerald Goertzel and since then is most commonly used in the detection of tones employed in Dual-tone Multi-frequency (DTMF) signaling \cite{b7}.
The working principle of the algorithm can be percieved as a second-order Infinite Impulse Response (IIR) filter (see \eqref{eqn:goertzel}) where the results $y[n-1]$ and $y[n-2]$ of the output $y[n]$
are stored in the states $Q_{1}$ and $Q_{2}$, so they are still available after the execution of the filter \cite{b8}.
\begin{equation}
    y[n] = x[n] + 2 \cdot cos(\omega_{g}) \cdot y[n-1] - y[n-2]
    \label{eqn:goertzel}
\end{equation}
\begin{figure}[h]
    \centering
    \resizebox{0.4\textwidth}{!}{
        \begin{tikzpicture}[auto, node distance=2cm]
            %Defining a cross set for the multiplication elements
            \tikzset{cross/.style={path picture={ 
            \draw[black]
            (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);
            }}}

@@

            \draw[->] (x.east) -- (sum1.west);
            \draw[->] (sum1.east) -- (y.west);
            \draw[->] (sum2.north) -- (sum1.south);
            \draw[->] (mult1.west) -- (sum2.east);
            \draw[->] (mult2.west) -- (0, -4) -- (sum2.south);
            \draw[->] (4, 0) -- (del1.north);
            \draw[->] (del1.south) -- (del2.north);
            \draw[->] (del2.south) -- (4,-4) -- (mult2.east);
            \draw[->] (4,-2) -- (mult1.east);
            \draw[->] (2, -1.25) -- (mult1.north);
            \draw[->] (2, -3.25) -- (mult2.north);
            \draw[->] (4, -2) -- (st1.west);
            \draw[->] (4, -4) -- (st2.west);
            \node at (2, -1) {$2 \cdot cos(\omega_{g})$};
            \node at (2, -3) {$-1$};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,0) {};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,-2) {};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,-4) {};
        \end{tikzpicture}
    }
    \caption{Block diagram of the Goertzel algorithm without the final calculation of the squared magnitude}
    \label{fig:goertzel_iir}
\end{figure}
\FloatBarrier\noindent
The system takes the input sequence $x[n]$ with length $N$ and is executed exactly $N$ times.
After the execution, the squared magnitude of $f_{a}$, the frequency to be analyzed inside $x[n]$, can be calculated using the values of the two states $Q_{1}$ and $Q_{2}$:
\begin{equation}
    |X(f_{a})|^2 = Q_{1}^2 + Q_{2}^2 - Q_{1} \cdot Q_{2} \cdot 2 \cdot \cos(\omega_{g})
\end{equation}
The constant $2 \cdot cos(\omega_{g})$, often called the "Goertzel coefficient", is calculated before the execution and used to adapt the algorithm to the frequency to be analyzed.
The Value of $\omega_{g}$ itself is calculated based on a integer number $k$ and $N$.
The integer number $k$ is a rounded fraction based on $N$, the frequency to be analyzed $f_{a}$, and the the sampling frequency $f_{s}$:
\begin{equation}
    k = \left\lfloor \frac{N \cdot f_{a}}{f_{s}} \right\rceil
\end{equation}
\begin{equation}
    w_g = \frac{2\pi \cdot k}{N}
\end{equation}

\par
Just as with the conventional FFT, the amount of needed calculations for executing the Goertzel algorithm is based on the number of samples.
In total, $N + 1$ multiplications and $2 \cdot N + 2$ additions are needed, which can be seen from the formulas for the IIR-part and the final calculation of the squared magnitude.
For example, a length of processing $N = 512$ samples requires in $513$ multiplications and $1026$ additions to calculate the squared magnitude of the desired frequency.
In comparison, the typical radix-2 FFT, needs $\frac{N}{2} \cdot \log_{2}(N)$ complex multiplications and $N \cdot \log_{2}(N)$ complex additions to be completed. With $N = 512$ this would result in $2034$ complex multiplications and $4608$ complex additions.
\par
With this simple example, it is already visible that the Goertzel algorithm can reduce the amount of needed resources compared to a full FFT, when only one or a small amount of frequency bins should be analyzed.
It is therefore used to observe the DCF77 signal's amplitude over time by being executed periodically.
As the gap of the encoded symbols is eiter \SI{100}{\milli\second} or \SI{200}{\milli\second} long, the duration of the analyzation set to \SI{10}{\milli\second} to provide a good tradeoff between resolution in time and a longest possible analyzation time
to leverage the DFT gain.
The square roots of the results are taken to finally calculate the magnitude of the DCF77 signal's amplitude and then passed to a detector to obtain the binary time signal.
Because of the duration of the analyzation being equal to \SI{10}{\milli\second}, the sampling rate after the Goertzel algorithm is equal to:
\begin{equation}
    f_{s,g} = \frac{1}{\SI{10}{\milli\second}} = \SI{100}{\hertz}
\end{equation}

\subsection{The detector}
The output of the Goertzel algorithm is already following the waveform of the binary time signal but still includes fluctuations due to noise and fading.
To reconstruct the original waveform of the binary time signal, it is processed inside a detector.
As the values of the Goertzel algorithm will follow the pattern of the signal's amplitude being switched between \SI{15}{\percent} and \SI{100}{\percent}, the detection can be done by comparing the value of the current sample to a threshold.
To compensate the changing of magnitude of the Goertzel algorithm's values that is caused by the current receiving situation, an Exponential Moving Average (EMA) is calculated by using \eqref{eqn:ema} and employed to create an adaptive threshold.
\begin{equation}
    y[t] = \alpha \cdot x[t] + (1 - \alpha) \cdot y[t-1]
    \label{eqn:ema}
\end{equation}
This part can also be represented as a block diagram:
\begin{figure}[h]
    \centering
    \resizebox{0.4\textwidth}{!}{
        \begin{tikzpicture}[auto, node distance=2cm]
            %Defining a cross set for the multiplication elements
            \tikzset{cross/.style={path picture={ 
            \draw[black]
            (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);
            }}}

@@

            \draw[->] (x.east) -- (mult1.west);
            \draw[->] (mult1.east) -- (sum1.west);
            \draw[->] (sum1.east) -- (y.west);
            \draw[->] (mult2.west) -- (1.5, -2) -- (sum1.south);
            \draw[->] (del1.south) -- (4.5, -2) -- (mult2.east);
            \draw[->] (4.5, 0) -- (del1.north);
            \draw[->] (0, 0.75) -- (mult1.north);
            \draw[->] (3, -1.25) -- (mult2.north);
            \node at (0, 1) {$\alpha$};
            \node at (3, -1) {$1 - \alpha$};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4.5,0) {};
        \end{tikzpicture}
    }
    \caption{Block diagram showing the calculation of the EMA}
    \label{fig:ema_iir}
\end{figure}
\FloatBarrier\noindent
The value of the smoothing factor $\alpha$ was calculated to take the samples of the last \SI{10}{\second} into account:
\begin{equation}
    \begin{split}
        \alpha &= \frac{2}{N + 1} = \frac{2}{\SI{10}{\second} \cdot f_{s,g} + 1}\\
        &= \frac{2}{\SI{10}{\second} \cdot \SI{100}{\hertz} + 1} \approx 0.001998
    \end{split}
\end{equation}
This provides a stable and slowly changing average, as it is putting more weight on the past samples and leaves minimal influence to the most recent sample.
The decision is then done by comparing the current sample to a threshold calculated by multiplying the current value of the average with $0.575$ (the middle bewteen \SI{15}{\percent} and \SI{100}{\percent}) and a scaling factor of $\frac{1}{0.8725}$.
The scaling factor is needed as the average percentage of time where the signal where is equal to \SI{100}{percent} is \SI{87.25}{\percent} (assuming an equal distribution between symbols representing a binary 1 and a 0).
If the value of the current sample is above this threshold, it is treated as a 1.
Otherwise it is treated as a 0.
The series of ones and zeros is then following the waveform shown in Fig. \ref{fig:dcf77_symbol_encoding} and can be used to decode the current time.

\section{MATLAB Simulation of the Receiver}
To verify and underline the working principle of the receiver, a simulation of the whole system was created using MATLAB.
The whole simulation was divided into five sections to allow for a clear structure:
\begin{enumerate}
    \item Simulation Parameter Section
    \item Signal Creation Section
    \item Goertzel Algorithm Section
    \item Detection Section
    \item FFT and Plot Section
\end{enumerate}

\section*{Acknowledgment}
The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\begin{thebibliography}{00}
\bibitem{b1} -, ``DCF77,'' 2024, Wikipedia entry. [Online]. Available: https://de.wikipedia.org/wiki/DCF77
\bibitem{b2} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77,'' PTB Mitteilungen, 2004, [Online]. Available: https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung\_4/4.4\_zeit\_und\_frequenz/4.42/dcf77.pdf
\bibitem{b3} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77: 1959-2009 und darüber hinaus,'' PTB Mitteilungen, 2009, [Online]. Available: https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung\_4/4.4\_zeit\_und\_frequenz/pdf/2009\_Bauch\_PTBM\_\_DCF77.pdf
\bibitem{b4} P. Hetzel, ``Time dissemination via the LF transmitter DCF77 using a pseudo-random phase-shift keying of the carrier,'' PTB Braunschweig, 1988, [Online]. Available: https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung\_4/4.4\_zeit\_und\_frequenz/pdf/5\_1988\_Hetzel\_-\_Proc\_EFTF\_88.pdf
\bibitem{b5} -, ``DCF77 webpage,'' 2024, Official page of the PTB [Online]. Available: https://www.ptb.de/cms/en/ptb/fachabteilungen/abt4/fb-44/ag-442/dissemination-of-legal-time/dcf77.html
\bibitem{b6} R. G. Lyons, ``Understanding Digital Signal Processing,'' Third Edition, Prentice Hall, 2010.
\bibitem{b9} J. Hoffmann, ``Sampling bandpass signals,'' 2021, Internet Article. [Online]. Available: https://www.dsprelated.com/showarticle/1407.php
\bibitem{b10} F. Zaplata and M. Kasal, ``SDR Implementation for DCF77,'', 2013, IEEE Paper. [Online]. Available: https://ieeexplore.ieee.org/document/6530943
\bibitem{b7} -, ``Goertzel algorithm,'' 2024, Wikipedia entry. [Online]. Available: https://en.wikipedia.org/wiki/Goertzel\_algorithm
\bibitem{b8} K. Banks, ``The Goertzel Algorithm,'' 2002, Internet Article. [Online]. Available: https://courses.cs.washington.edu/courses/cse466/12au/calendar/Goertzel-EETimes.pdf
\end{thebibliography}
\end{document}
